A methodology for calculating active force-time functions from broadband piezoelectric sensors

### 2020-05-01 Article_file_Source_reconstruction.tex
Updates from Overleaf

\begin{abstract}

\end{abstract}

\begin{keyword}
Acoustic emissions,  piezoelectric sensors, absolute source calibration 
\end{keyword}
\end{frontmatter}

@@

While more complex sources have been considered to model earthquakes, such as the single and double couple source \citet{Sato1972}, we believe the original problem studied by \citet{Lamb1904} -- elastic displacements resulting from a point force in/on a half space -- is sufficiently complex for capturing the behavior of the piezoelectric actuator depicted in Figure \ref{fig2}(a). In Figure \ref{fig2}(a), we show the cross-section A-A from Figure \ref{fig1}(b). The source is produced by pulsing the piezoelectric actuator on the top surface using the experimental facilities described in Figure \ref{fig1}. The impulsive actuation of the PZT crystal on the steel plate produces a rapid transient source causing elastodynamic stress waves to propagate outwards. In Figure \ref{fig2}, the coloured region represents a synthetic representation of the velocity wave-field and is only provided as a reference and is not representative of the source produced from the PZT actuator (which is to be determined later). For this example, a snapshot of the velocity wave-field, produced from a capillary fracture is shown from numerical finite element simulations in ABAQUS \citep[adapted from fig. 2][]{Selvadurai2019}.

Properties of the source creating the stress waves is important. \citet{Breckenridge1990} states that the ideal source is infinitesimal in size, has a force-time function of a step or delta function, and generates large amplitudes. He performed important study of a variety of sources including pencil lead fracture, capillary fracture, capacitive transducer, conical transducer, ball impact, spark, and high explosive, noting that each type is imperfect and has its trade offs. In this study, we look to aim to understand the PZT actuator (similar to conical transducer) which we can leverage to non-invasively study the \textit{in situ} absolute calibration conditions inside a less accessible experimental facility (e.g. trixaial pressure cell). To model the force-time function produced conical sensor, we first simplify the sequence of events and their components that link the source to the recorded signal.

In Figure \ref{fig2}(a), we show the five components of ``toolchain'' that is used to link the source to the measured signal. The PZT actuator is used to produce a (\textit{1}) \textbf{rapid transient force} $f_{j}$ on the top surface at a point $\mathbf{\xi}$ of the steel platen at a delayed time $\tau$. Elastic stress waves propagate through the plate and the theoretical elastic displacement field throughout the body can be computed using the (\textit{2}) \textbf{Green's function} $G_{kj}$. The Green's functions map the (\textit{3})\textbf{ elastic displacements} $u_{k}$ at $\mathbf{x}$ and time $t$ from the force at $\mathbf{\xi}$ and time $\tau$. A passive sensor, on the bottom surface at point $\mathbf{x}$ and attempts to measure the theoretical displacements $u_{k}$. Unfortunately, the passive sensors always will distort the (\textit{4}) \textbf{measured signal}  $s$ according to its (\textit{5}) \textbf{instrument response} $i_{k}$ at position $\mathbf{x}$ and time $t$.

@@

A total of 13 sensor are used and the spectral characteristics of these sensors have been well examined in the literature using the absolute calibration methods \citep{Glaser1998, McLaskey2010, McLaskey2012, Selvadurai2015, Selvadurai2019, Wu2020}. \citet{Glaser1998} provides a detailed account to sensor attributes used by KRN Services, including a the small field emission transistor (FET)-based Sanyo (2SK715W) preamplifier internal to the sensor casing to improve impedance matching.  These sensors have shown flat displacements-responses between 100 kHz to 1 MHz \citep{McLaskey2012, Selvadurai2019}.  At longer wavelengths between 1 kHz to 50 kHz these sensors have been found to behave more like accelerometer \citep{Wu2020}.

@@

\noindent where $I_{k}$, $S$, $F_{j}$ and $G_{kj}$ are calculated from the temporal Fourier transforms of the $i_{k}$, $s$, $f_{j}$ and $g_{kj}$, respectively.  In a concerted study, we have predetermined the instrument responses for the exact same sensor configuration shown in Figure \ref{fig1}.  However, \citet{Wu2020} produced a known force-time function by dropping small steel balls of different diameters onto the surface from a known height at the identical location that will be used for the PZT actuator later. According to Hertzian contact theory, it is possible to assume a source time function that is a function of the material properties of the ball and steel platen ($E, \nu, \rho$), the radius of ball, and the drop height. This method has been popularized over the recent years due to a highly reliable and repeatable force-time function $f$ produced from ball impact and the ability to absolutely calibrate the acoustic emission sensors over a wide spectral range \citep{Breckenridge1990, McLaskey2010, McLaskey2012, McLaskey2015}.

For the identical configuration of sensors shown in Figure \ref{fig1}(b), \citet{Wu2020} has dropped a range of small stainless steel (404) balls with diameters ranging from $d_{ball}$ = 0.3, 0.35, 0.4, 0.5, 0.6, 0.7, 0.8, 1.0, 1.5, 2.0 and 2.5 mm. The balls had elastic modulii $E_{ball}$ = 210 GPa and Poisson ratio $\nu_{ball}$ = 0.303, which allow them to calcaute the specific force-time function from Hertzian contact theory \citep[see also e.g.][]{McLaskey2010, McLaskey2012}. Once the force-time function is know, we use equation \eqref{eq5}, now with $F_{ball}\left( \mathbf{\xi}, \omega \right)$, which represents the Fourier transform of the ball drop, to calculate the broadband instrument response $I_{k}(\omega)$ from 1 kHz to 1 MHz.  Calculating the lower frequency instrument response (i.e. 1 kHz to 50 kHz ) is not straightforward and the readers should also consult \citet{Wu2018} for a summary of methods. Figure \ref{fig6}(a) shows the full instrument response for all the passive sensors shown in Figure \ref{fig1}(b). The color coding related to the sensors epicentral distance from the ball drop source. In this study, we are only interested in the higher frequency (100 kHz to 1 MHz) sensor behavior for the reconstruction of the PZT actuator force time function. This range is shown in Figure \ref{fig6}(b). We see that the instrument response is insensitive to the take-off angle. This is important to the analysis and will be addressed in the more detail later. This insensitivity has also been acknowledged in the past for the same KRNPC-BB conical-style sensors used here \citep{Goodfellow2015, Selvadurai2019}.  Results from this calibration campaign are summarized and more details have been given in \citet{Wu2020}.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 1.0]{FIG6.pdf} 
\caption{Summary of the calibration experiments performed by \citet{Wu2020} to determine the instrument response for ball drops, performed \textit{a priori}, in the same location as where our PZT actuator was later placed. \textbf{(a)} The full range of calibrated frequencies (1 kHz to 1 MHz) is shown. \textbf{(b)} For our study, we only consider high-frequency aspect and therefore will only use the instrument responses between components 100 kHz to 1 MHz.  Colors are indicative of the sensors location as defined in Figure \ref{fig1}(b). For all the details please consult \citet{Wu2020}. }
	\label{fig6} 
\end{figure}

\subsection{Spectral Deconvolution of the Problem}
\label{Spec_deconv}
Having the specific broadband instrument response, solved using the well-understood ball drop source for our exact sensor-source layout, is paramount to accurately model the force-time function of the PZT actuator described by equation \eqref{eq4b}.   If we now consider the original problem in Figure \ref{fig2}, we can start with the recorded signal produced from the PZT actuator and use the instrument response to determined the mechanical vibration.  In the Fourier domain, we can rearrange the left hand side of equation \eqref{eq3} to estimate the spectral response of the theoretical vibration $U_{k}(\mathbf{x},t)$ caused by the transient PZT actuator:

@@

\section{Results}
\label{results}
\subsection{Time domain analysis}
A total of 1000 pulses of the PZT actuator were performed but for our results we show a single pulse to show the methodology that follows equations \eqref{eq1} to \eqref{eq6b}.  The source was highly repeatable in at the high-frequencies we investigated.  From the 1000 tests, the average  Pearson correlation coefficient from the stacked mean was $R$ = \textbf{0.95}. In Figure \ref{fig3}(a) we show measurements on all passive sensors PZT sensors depicted in Figure \ref{fig1}(b) for a random pulse with color indicating different epicentral distances. For our analysis, we only the first 50 $\mu$s of signal since we do not want to pollute the data with reflections from the boundary of the platen that would occur if longer signals were analyzed. The sampling rate was $F_{s}$ = 20 MHz making each trace 1000 samples. Since boundary reflections are not considered, we can approximate that the plate is in fact semi-infinite. Under this assumption we are able to solve for the Green's functions using the generalized ray theory code provided by \citet{Hsu1985} and modified for MATLAB by \citet{McLaskey2012}.  Figure \ref{fig3}(b) shows the Green's functions at different epicentral locations, moving out from the source at $x$ = 0 mm. We call these approximations the empirical Green's function (EGF) since it is a numerical approximation of the true Green's function.  The true Green's function can be calculated by replacing the force-time function in equatoin \eqref{eq1} with the a unit impulse (Dirac delta) function $\delta_{j}(\mathbf{\xi}, \tau)$. In Figure \ref{fig3}(b) we also show the wave phase arrivals (P, S, PPP, PPPPP) from the numerical models.

@@

@@

\subsection{Frequency domain analysis}
In this section we perform the discrete Fourier transform of the measured and theoretical traces in Figure \ref{fig3}.  For clarity we only show the results from the sensor A3 shown in Figure \ref{fig5} to illustrate the steps taken to solve for the scaling function $A_{0}$. To minimize the uncertainty that can occur from non-uniform spacing of samples of the Fourier frequencies linear scale, we average over multiple bins while evenly spacing our samples on the logarithmic scale.  We re-sample all DFT on the with 41 samples from 100 kHz to 1000 kHz.  While this might cause some averaging uncertainties between neighboring frequencies, this does not affect our final estimates of the scaling factor  $A_{0}$ in equation \eqref{eq4c} since, as we will see, we take estimate the scaling factor $A_{0}$ as a single average over the full range of frequencies.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 1]{FIG7.pdf} 
\caption{\textbf{(a)} Discrete Fourier transform (DFT) of the measured waveform $S$ (green line) and the instrument response from sensor $I_{k}$ (black line) at A3 between 100 kHz to 1000 kHz. \textbf{(b)} Spectral estimates of the ground displacements $U_{k}$ (gray line) obtained from spectral division using equation \eqref{eq6}. \textbf{(c)} We show the DFT of the normalized force-time function $F_{j}^{*}$ (blue line) and the empirical Green's function $G_{kj}$ (black line) . \textbf{(d)} Spectral estimates of the corrected force time function $F$ calculated from the spectral estimates of the signal, instrument response, Green's function, normalized force time function. The spectral mean from between 100 kHz to 1000 kHz is shown.}
	\label{fig7} 
\end{figure}

@@

Using the calculations highlighted in Figure \ref{fig7}, we use the inverse Discrete Fourier transform (IDFT) to return the force-time function into the time domain. Firstly, due to the fact that the frequency response is relatively flat over the 100 kHz to 1000 kHz range (Figure \ref{fig7}(d)) we use the mean spectral value to scale the spectrum of normalized force-time function $F^{*}$ (blue line in Figure \ref{fig7}(c)) prior to performing the IDFT. We choose to scale the normalized source time function since, due to properties of the IDFT, we require frequencies outside the range we are able to estimate the spectral calculations shown in Figure \ref{fig7}.  We note that lower frequencies cannot be studied due to the plate finite geometry that limits our analysis of the signal to 50 $\mu s$.  In Figure \ref{fig9}(a) we show the broadband DFT of the normalized force-time function (blue line).  We scale this with the scaling factor from Figure \ref{fig7}(d) to obtain the re-scaled force-time function in the Fourier domain $F$ (black line). We have also applied a 4-pole bandpass butterworth filter (red line) between 100 kHz and 1000 kHz, which will be used in our time domain comparison between synthetic estimates to the filtered and unfiltered signals. Figure \ref{fig9}(b) shows the IDFT to our estimates of the normalized (blue), unfiltered (black) and filtered (red) signals.  The corrected force-time function in the time domain estimates that the peak force of $\sim$ 3.4 N. This estimate fall within a reasonable range of estimated force-time function for conical sensor estimate by \citet{Breckenridge1990} $\sim $ 1.2 N for a transient step function over 1 $\mu s$ with an amplitude of 100 V. We note that if the systems studied by \citet{Breckenridge1990} was linear, for a 200 V we would expect a 2.4 N step.  While there are discrepancies, the exact details of the conical PZT (geometry, casing, backing mass, damping material, etc.) are not given so it is difficult to perfectly characterize the discrepancies. Moreover, the voltage-time function that excited the PZT actuator differs in each study; we employ a pulse-discharge excitation, whereas they employed a step-hold.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 1]{FIG9.pdf} 
\caption{\textbf{(a)}  }
	\label{fig9} 
\end{figure}

\subsection{Synthetic reconstructing}

@@

\clearpage

\section*{Acknowledgement}
T. M\"orgeli

@@

\section*{Supplemental information}

@@

\end{document}
\endinput

>>>>

\begin{abstract}
We preform active source tests using a piezoelectric (PZT) actuator that was excited by a rapid transient high voltage pulse (HVP). The crystal geometry of the actuator was a conical frustum where the tip was placed centrally on top of a thick steel plate. We studied the elastodynamic stress waves  (between 100 kHz to 1 MHz) produced by the source, whereby the HVP pulse was converted to a rapid transient force-time function ($f_{j}$) via the piezoelectric effect.  The source produced mechanical vibrations ($u_{k}$) throughout the sample and was measured ($s$) using an array of 13 passive conical PZT receivers, at known locations on the bottom of the steel plate. Dynamic Green's functions ($G_{kj}$) were calculated using generalized ray theory and the known instrument responses ($I_{k}$) were pre-calculated from a series of ball drop experiments for all source-receiver pairs. Using these parameters, we were able to reconstruct the force-time function of the PZT actuator. As predicted from the literature, the actuator behaved as an under-damped oscillator with a peak loads, decay rate and undamped resonant frequency of $A_{0}$ = 3.3 $\pm$ 0.8 N, $\Gamma$ = 0.7 MHz and $\Lambda$ = 4.7 MHz, respectively.  Both the decay rate and resonant frequency were determined as by minimizing the error between theoretical displacements for this force-time function over all 13 receiver measurements. Understanding calibrated force-time functions from PZT actuators can help to develop time-dependent and nondestructive characterization of many acoustic emission system components, such as the receiving transducers or the transfer media itself, in non-accessible laboratory environments. 
\end{abstract}

\begin{keyword}
Laboratory seismology, acoustic emissions,  piezoelectric sensors, absolute source calibration, 
\end{keyword}
\end{frontmatter}

@@

While more complex sources have been considered to model earthquakes, such as the single and double couple source, we believe the original problem studied by \citet{Lamb1904} -- elastic displacements resulting from a point force in/on a half space -- is sufficiently complex for capturing the behavior of the piezoelectric actuator depicted in Figure \ref{fig2}(a). In Figure \ref{fig2}(a), we show the cross-section A-A from Figure \ref{fig1}(b). The source is produced by pulsing the piezoelectric actuator on the top surface using the experimental facilities described in Figure \ref{fig1}. The impulsive actuation of the PZT crystal on the steel plate produces a rapid transient source causing elastodynamic stress waves to propagate outwards. In Figure \ref{fig2}, the coloured region represents a synthetic representation of the velocity wave-field and is only provided as a reference and is not representative of the source produced from the PZT actuator (which is to be determined later). For this example, a snapshot of the velocity wave-field, produced from a capillary fracture is shown from numerical finite element simulations in ABAQUS \citep[adapted from fig. 2][]{Selvadurai2019}.

Properties of the source creating the stress waves is important. \citet{Breckenridge1990} states that the ideal source is infinitesimal in size, has a force-time function of a step or delta function, and generates large amplitudes. He performed important study of a variety of sources including pencil lead fracture, capillary fracture, capacitive transducer, conical transducer, ball impact, spark, and high explosive, noting that each type is imperfect and has its trade offs. In this study, we look to aim to understand the PZT actuator (similar to conical transducer) which we can leverage to non-invasively study the \textit{in situ} absolute calibration conditions inside a less accessible experimental facility (e.g. triaxial pressure cell). To model the force-time function produced conical sensor, we first simplify the sequence of events and their components that link the source to the recorded signal.

In Figure \ref{fig2}(a), we show the five components of ``toolchain'' that is used to link the source to the measured signal. The PZT actuator is used to produce a (\textit{1}) \textbf{rapid transient force} $f_{j}$ on the top surface at a point $\mathbf{\xi}$ of the steel platen at a delayed time $\tau$. Elastic stress waves propagate through the plate and the theoretical elastic displacement field throughout the body can be computed using the (\textit{2}) \textbf{Green's function} $G_{kj}$. The Green's functions map the (\textit{3}) \textbf{mechanical vibrations} $u_{k}$ at $\mathbf{x}$ and time $t$ from the force at $\mathbf{\xi}$ and time $\tau$. A passive sensor, on the bottom surface at point $\mathbf{x}$ and attempts to measure the theoretical displacements $u_{k}$. Unfortunately, the passive sensors always will distort the (\textit{4}) \textbf{measured signal}  $s$ according to its (\textit{5}) \textbf{instrument response} $i_{k}$ at position $\mathbf{x}$ and time $t$.

@@

A total of 13 sensor are used and the spectral characteristics of these sensors have been well examined in the literature using the absolute calibration methods \citep{Glaser1998, McLaskey2010, McLaskey2012, Selvadurai2015, Selvadurai2019, Wu}. \citet{Glaser1998} provides a detailed account to sensor attributes used by KRN Services, including a the small field emission transistor (FET)-based Sanyo (2SK715W) preamplifier internal to the sensor casing to improve impedance matching.  These sensors have shown flat displacements-responses between 100 kHz to 1 MHz \citep{McLaskey2012, Selvadurai2019}.  At longer wavelengths between 1 kHz to 50 kHz these sensors have been found to behave more like accelerometer \citep{Wu}.

@@

\noindent where $I_{k}$, $S$, $F_{j}$ and $G_{kj}$ are calculated from the temporal Fourier transforms of the $i_{k}$, $s$, $f_{j}$ and $g_{kj}$, respectively.  In a concerted study, we have predetermined the instrument responses for the exact same sensor configuration shown in Figure \ref{fig1}.  However, \citet{Wu} produced a known force-time function by dropping small steel balls of different diameters onto the surface from a known height at the identical location that will be used for the PZT actuator later. According to Hertzian contact theory, it is possible to assume a source time function that is a function of the material properties of the ball and steel platen ($E, \nu, \rho$), the radius of ball, and the drop height. This method has been popularized over the recent years due to a highly reliable and repeatable force-time function $f$ produced from ball impact and the ability to absolutely calibrate the acoustic emission sensors over a wide spectral range \citep{Breckenridge1990, McLaskey2010, McLaskey2012, McLaskey2015}.

For the identical configuration of sensors shown in Figure \ref{fig1}(b), \citet{Wu} has dropped a range of small stainless steel (404) balls with diameters ranging from $d_{ball}$ = 0.3, 0.35, 0.4, 0.5, 0.6, 0.7, 0.8, 1.0, 1.5, 2.0 and 2.5 mm. The balls had elastic modulii $E_{ball}$ = 210 GPa and Poisson ratio $\nu_{ball}$ = 0.303, which allow them to calcaute the specific force-time function from Hertzian contact theory \citep[see also e.g.][]{McLaskey2010, McLaskey2012}. Once the force-time function is know, we use equation \eqref{eq5}, now with $F_{ball}\left( \mathbf{\xi}, \omega \right)$, which represents the Fourier transform of the ball drop, to calculate the broadband instrument response $I_{k}(\omega)$ from 1 kHz to 1 MHz.  Calculating the lower frequency instrument response (i.e. 1 kHz to 50 kHz ) is not straightforward and the readers should also consult \citet{Wu2018} for a summary of methods. Figure \ref{fig6}(a) shows the full instrument response for all the passive sensors shown in Figure \ref{fig1}(b). The color coding related to the sensors epicentral distance from the ball drop source. In this study, we are only interested in the higher frequency (100 kHz to 1 MHz) sensor behavior for the reconstruction of the PZT actuator force time function. This range is shown in Figure \ref{fig6}(b). We see that the instrument response is insensitive to the take-off angle. This is important to the analysis and will be addressed in the more detail later. This insensitivity has also been acknowledged in the past for the same KRNPC-BB conical-style sensors used here \citep{Goodfellow2015, Selvadurai2019}.  Results from this calibration campaign are summarized and more details have been given in \citet{Wu}.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 1.0]{FIG6.pdf} 
\caption{Summary of the calibration experiments performed by \citet{Wu} to determine the instrument response for ball drops, performed \textit{a priori}, in the same location as where our PZT actuator was later placed. \textbf{(a)} The full range of calibrated frequencies (1 kHz to 1 MHz) is shown. \textbf{(b)} For our study, we only consider high-frequency aspect and therefore will only use the instrument responses between components 100 kHz to 1 MHz.  Colors are indicative of the sensors location as defined in Figure \ref{fig1}(b). For all the details please consult \citet{Wu}. }
	\label{fig6} 
\end{figure}

\subsection{Spectral Deconvolution}
\label{Spec_deconv}
Having the specific broadband instrument response, solved using the well-understood ball drop source for our exact sensor-source layout, is paramount to accurately model the force-time function of the PZT actuator described by equation \eqref{eq4b}.   If we now consider the original problem in Figure \ref{fig2}, we can start with the recorded signal produced from the PZT actuator and use the instrument response to determined the mechanical vibration.  In the Fourier domain, we can rearrange the left hand side of equation \eqref{eq3} to estimate the spectral response of the theoretical vibration $U_{k}(\mathbf{x},t)$ caused by the transient PZT actuator:

@@

\section{Results}
\label{results}
\subsection{Time domain analysis}
To test for source repeatability, we looked a a sweep of 15 pulses of the PZT actuator were performed.  The source was highly repeatable in at the high-frequencies we investigated.  From the 13 tests, the Pearson correlation coefficient (PCC) was analysed with respected to the stacked mean. We found that the average PCC = 0.9646 $\pm$ 0.03 over all the sensors for the 50 $\mu s$ time window analysed. We believe this to be representative of a repeatable source with respect to waveform similarities. Our final results will take these deviations into consideration; however, for clarity the results shown in this Section will focus on a single actuator pulse to highlight the methodology described in equations \eqref{eq1} to \eqref{eq6b}.

@@

In Figure \ref{fig3}(a) we show measurements on all passive sensors PZT sensors depicted in Figure \ref{fig1}(b) for a random pulse with color indicating different epicentral distances. For our analysis, we only the first 50 $\mu$s of signal since we do not want to pollute the data with reflections from the boundary of the platen that would occur if longer signals were analyzed. The sampling rate was $F_{s}$ = 20 MHz. Since boundary reflections are not considered, we can approximate that the plate is in fact semi-infinite. Under this assumption we are able to solve for the Green's functions using the generalized ray theory code provided by \citet{Hsu1985} and modified for MATLAB by \citet{McLaskey2012}.  Figure \ref{fig3}(b) shows the Green's functions at different epicentral locations, moving out from the source at $x$ = 0 mm. We call these approximations the empirical Green's function (EGF) since it is a numerical approximation of the true Green's function.  The true Green's function can be calculated by replacing the force-time function in equatoin \eqref{eq1} with the a unit impulse (Dirac delta) function $\delta_{j}(\mathbf{\xi}, \tau)$. In Figure \ref{fig3}(b) we also show the wave phase arrivals (P, S, PPP, PPPPP) from the numerical models.

@@

\subsection{Frequency domain analysis}
In this section we perform the discrete Fourier transform of the measured and theoretical traces in Figure \ref{fig3}.  For clarity we only show the results from the sensor A3 shown in Figure \ref{fig5} to illustrate the steps taken to solve for the scaling function $A_{0}$. To minimize the uncertainty that can occur from non-uniform spacing of samples of the Fourier frequencies linear scale, we average over multiple bins while evenly spacing our samples on the logarithmic scale.  We re-sample all DFT on the with 41 samples from 100 kHz to 1000 kHz.  While this might cause some averaging uncertainties between neighboring frequencies, this does not affect our final estimates of the scaling factor  $A_{0}$ in equation \eqref{eq6c} since, as we will see, we take estimate the scaling factor $A_{0}$ as a single average over the full range of frequencies.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 0.9]{FIG7.pdf} 
\caption{\textbf{(a)} Discrete Fourier transform (DFT) of the measured waveform $S$ (green line) and the instrument response from sensor $I_{k}$ (black line) at A3 between 100 kHz to 1000 kHz. \textbf{(b)} Spectral estimates of the ground displacements $U_{k}$ (gray line) obtained from spectral division using equation \eqref{eq6}. \textbf{(c)} We show the DFT of the normalized force-time function $F_{j}^{*}$ (blue line) and the empirical Green's function $G_{kj}$ (black line) . \textbf{(d)} Spectral estimates of the corrected force time function $F$ calculated from the spectral estimates of the signal, instrument response, Green's function, normalized force time function. The spectral mean from between 100 kHz to 1000 kHz is shown.}
	\label{fig7} 
\end{figure}

@@

Using the calculations highlighted in Figure \ref{fig7}, we use the inverse Discrete Fourier transform (IDFT) to return the force-time function into the time domain. Firstly, due to the fact that the frequency response is relatively flat over the 100 kHz to 1000 kHz range (Figure \ref{fig7}(d)) we use the mean spectral value to scale the spectrum of normalized force-time function $F^{*}$ (blue line in Figure \ref{fig7}(c)) prior to performing the IDFT. We choose to scale the normalized source time function since, due to properties of the IDFT, we require frequencies outside the range we are able to estimate the spectral calculations shown in Figure \ref{fig7}.  We note that lower frequencies cannot be studied due to the plate finite geometry that limits our analysis of the signal to 50 $\mu s$.

Figure \ref{fig9}(a) shows the all Fourier frequencies of our synthetic normalized force-time function $F^{*}$ from 20 kHz to 5 MHz (blue). Using the scaling factor from Figure \ref{fig7}(d), we can adjust our estimates that corrects our force-time function $F$ (black line) with calibration estimates from the instrument response determined form the ball drop. We have also applied a 4-pole bandpass butterworth filter $filt(F)$ (red line) between 100 kHz and 1000 kHz, which will be used in our time domain comparison between synthetic estimates to the filtered and unfiltered signals. Figure \ref{fig9}(b) shows the IDFT to our estimates of the normalized (blue), unfiltered (black) and filtered (red) force-time functions.  The corrected force-time function (black) in the time domain estimates that the peak force of $\sim$ 3.4 N.

\begin{figure}[ht]
     	\centering
\includegraphics[scale=0.9]{FIG9.pdf} 
\caption{\textbf{(a)} Discrete Fourier transform (DFT) of the normalized $F^{*}$ (blue line), scaled $F$ (black line) and bandpass filtered $filt(F)$ (red line) of the force time function shown in Figure \ref{fig2b}(d). We show scaling factor of 3.4 N/Hz determined from the spectral analysis. \textbf{(b)} Inverse discrete Fourier transform (IDFT) of the signals in (a). }
	\label{fig9} 
\end{figure}
\subsection{Effects of receiver location on force-time estimates}

To better asses the variations in our estimates of $A_{0}$ with the spatial locations of the receivers we looked at results from all sensors for the 15 pulses. Figure \ref{fig9a}(a) shows the estimates of $A_{0}$ with takeoff angle $\theta$. The takeoff angle for this setup is depicted in Figure \ref{fig2}.  Piezoelectric transducers used in the laboratory have been known to be sensitive to incident angle of the incoming waves \citep{Goodfellow2015}.  This is also seen at the mining scale scale \citep{Kwaitek2011}. We see that there is a slight decrease in estimates of $A_{0}$ with increasing  $\theta$, with an average value of 2.3 N.

Figure \ref{fig9a}(b) looks at dependence of $A_{0}$ on the total distance travelled (hypocentral distance).  Attenuation is the intrinsic loss in energy associated with a cycle of the wave for example by the internal friction of the material.  By fitting an exponential decay curve we can estimate the materials quality factor $Q$ -- a metric to describe the losses per cycle.  Our fit finds $A^{*}$ = 4.7 N and $\alpha$ = 0.01. Using these parameters we found the average $Q$ = 20,470 between 100 kHz to 1000 kHz, which is acceptable for steel (see Supplemental S2 for more details). It is impossible to perfectly determine whether the loss of sensitivity is related to the material property ($Q$) or a sensor characteristics due to the linear dependency of the hypocentral distance on $\theta$, however; material attenuation can explain this decrease with some degree of certainty. Taking into account attenuation, the value of the force at the source was $A^{*}$ = 4.7 N. This estimate falls within a reasonable range of estimated force-time function for conical sensor by \citet{Breckenridge1990} $\sim $ 1.2 N for a transient step function over 1 $\mu s$ with an amplitude of 100 V. If the systems studied by \citet{Breckenridge1990} was linear, we expect that 200 V we would result in a 2.4 N step.  This discrepancy may results from properties of the system (PZT geometry, casing, backing mass, damping material, electronics, data acquisition system, etc.), which is not identical to our making it difficult to perfectly characterize why they differ discrepancies. Moreover, the voltage-time function that excited the PZT actuator differs in each study; we employ a pulse-discharge excitation, whereas they employed a step-hold. However, we feel that the similarity in the order of $A_{0}$ is a positive results.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 1]{FIG9a.pdf} 
\caption{Relationships between spectral estimates of $A_{0}$ and the \textbf{(a)} takeoff angle ($\theta$) and \textbf{(b)} hypocentral distance ($R$). An exponential fit was used to determine the quality factor $Q$ (see Supplemental Section S2 for more details)  }
	\label{fig9a} 
\end{figure}
\subsection{Synthetic reconstructing}
Figure \ref{fig10} shows a comparison of the theoretical ground displacements for the estimated (a) unfiltered and (b) bandpass filtered force-time functions from Figure \ref{fig9}(b).  To calculate the theoretical displacements we convolve the force-time function with the Green's function following equation \eqref{eq2}. We see that the theoretical vibrations match the theory quite we on all sensors. This further enforces our hypothesis that the decrease in force estimates on sensors at higher takeoff angles was not a sensor sensitive issue but due to material attenuation. Figure \ref{fig10} is a rewarding result from this study and highlights the potential fully understanding the force-time function produced by our PZT actuator and HVP unit.

@@

\section{Conclusions}
We have followed the recently popularized methodology for sensor calibration coupled reasonable understanding of the wave propagation problem in a semi-infinite, homogeneous elastic half space. By fully characterizing the individual components of our system, we show that it is possible to better understand the force-produced by a conical PZT actuator when excited by a rapid high-voltage transient. While the results presented here are unique to our system, the methodology and insight into the characteristic shape of the force-time function can be easily reproduced in other acoustic emission system. Fully understanding how system responses change during an experiment is critical to properly interpret seismicity in a quantitative manner. Our approach for time-dependent updates to calibration parameters but also can provide intrinsic and quantitative understanding of changes in the transfer media.  
\section*{Acknowledgement}
T. M\"orgeli

@@

\clearpage
\section*{Supplemental information}

\subsection{Material Properties of PZT-5A (Navy II)}

@@

\clearpage
\subsection{S2 Quality Factor}
We estimate the quality factor in steel $Q_{steel}$ from estimates of the force magnitude $A_{0}$ at different hypocentral distances. This follow typical theory on attenuation used to model the damping of seismic waves over distance due to internal intrinsic losses (e.g. friciton) in the material that decreased the energy per cycle\citep[e.g.][]{Barton2007}. From Figure \ref{fig9}(b) we have fit the exponential decay over distance with $A^{*}$ = 4.7 N and $\alpha$ = 0.01 m$^{-1}$. The quality factor depends on the frequencies of the propagating wave and can be expressed as a function as follows

\begin{equation}
    \label{eqS2}
Q = \frac{\pi f}{V_{P}\alpha}.
\end{equation}

\noindent We can plot the quality factor over the frequency range 100 kHz to 1 MHz investigated in this study for $\alpha$ = 0.01 m$^{-1}$. Typical values of $Q_{steel} > $ 10,000 in the literature \citep{WaitingRUI}. Figure \ref{figS1} shows that at $f$ =100 kHz $Q$ = 5193, a relatively low estimate, and at $f$ = 1 MHz $Q$ = 51,935, which is a higher estimate for steel. Assume a constant attenuation quality factor of the range of 100 kHz to 1 MHz we obtain a average $Q$ = 20,470.

\begin{figure}[ht]
     	\centering
\includegraphics[scale= 1]{FIGS1.pdf} 
\caption{\textbf{(a)}  }
	\label{figS1} 
\end{figure}

\end{document}

\endinput
